---
title: "Spotify"
output:
  html_notebook:
    css: style.css
runtime: shiny
---

I have consistently been interested in assessing music information in R: see. for example my - somewhat dormant -[charts dashboard](https://mytinyshinys.shinyapps.io/musicChartsFlexDB/)

A recent, excellent, [blog post by RCharlie](http://rcharlie.com/2017-02-16-fitteR-happieR/#comment-3169008481) featuring the Spotify and Genius APIs to determine a 'Gloom Index' for Radiohead tracks piqued my interest. It also incorporates some illuminating code using the purrr package

My intention, initially at least, is just to use the Spotify, i.e. musical section of his work  and make it a Shiny app which will allow the full catalogue  to be accessed. I extend it to a few different outputs, particularly the ability to link to youtube videos of chosen tracks. 
 Much of the code is a complete steal from RCharlie and mine could definitely do with some enhancements.    
 
 *Please note that if you plan to use the code do obtain your own [client authorization](https://developer.spotify.com/web-api/authorization-guide/)*
 
 
 ---

```{r, setup, warning= FALSE, echo=TRUE, error=FALSE, message= FALSE}
knitr::opts_chunk$set(eval=TRUE, echo=TRUE, error=FALSE, message= FALSE)

#Load required libraries

library(httr)
library(stringr)
library(tidyverse)
library(lubridate)
library(rvest)
library(tidytext)
library(scales)
library(RColorBrewer)
library(highcharter)
library(DT)
library(shiny)
library(vembedr)
library(plotly)

```

```{r sp_authorization,  echo=FALSE}
client_id <- 'a749814f307b45fc9144e5ab6a0436e1'
  client_secret <- '7beb57b32692463589c0491d9e35f60d'
  access_token <- POST('https://accounts.spotify.com/api/token',
                       accept_json(), authenticate(client_id, client_secret),
                       body = list(grant_type='client_credentials'),
                       encode = 'form', httr::config(http_version=2)) %>% content %>% .$access_token


```


Firstly, enter the artist of interest and click the first button

```{r artist, warning=FALSE}

# Select Artist  

textInput("artist","Enter Artist or first few letters",value="", placeholder="e.g. Laura Nyro" )
actionButton(inputId="goArtist",label="Get Artist Selection")

hr()
  
  artistData <- eventReactive(input$goArtist,{

    req(input$artist)
     req(input$artist)
  # Search Spotify API for artist name
  res <- GET('https://api.spotify.com/v1/search', query = list(q = input$artist, type = 'artist')) %>%
    content %>% .$artists %>% .$items

  # Clean response and combine all returned artists into a dataframe
  artists <- map_df(seq_len(length(res)), function(x) {
    list(
      artist_name = res[[x]]$name,
      artist_uri = str_replace(res[[x]]$uri, 'spotify:artist:', ''), # remove meta info from the uri string
      artist_img = ifelse(length(res[[x]]$images) > 0, res[[x]]$images[[1]]$url, NA)
    )
  })

   return(artists)
  })
  
  output$artists <- DT::renderDataTable({
   artistData() %>%
      #select(artist=artist_name,artist_uri,artist_img) %>% 
      select(artist=artist_name) %>% 
  DT::datatable(class='compact stripe hover row-border order-column',
                selection="multiple",rownames=FALSE,options= list(paging = FALSE, searching = FALSE,info=FALSE))
  })
br()
  DT::dataTableOutput("artists")
br()
actionButton(inputId="goArtists",label="Get Exact Artist(s)")


  
```  


Now select row(s) from the resultant table. For instance, you will probably not want karaoke or tribute bands but may or may not wish to include collaborations

---


```{r albums, warning=FALSE}



#Select Albums


 allArtistsData <- eventReactive(input$goArtists,{

   req(input$artists_rows_selected)
  s = as.integer(input$artists_rows_selected)

 
  ## should purrr this, presumably
  for (i in seq_along(s)) {

    j <- s[i]
  
 albums <- GET(paste0('https://api.spotify.com/v1/artists/', artistData()$artist_uri[j],'/albums')) %>% content


album_info <-  map_df(1:length(albums$items), function(x) {
    tmp <- albums$items[[x]]
    
    # Make sure the album_type is not "single"
    if (tmp$album_type == 'album') {
      data.frame(album_uri = str_replace(tmp$uri, 'spotify:album:', ''),
                 album_name = str_replace_all(tmp$name, '\'', ''),
                 album_img = albums$items[[x]]$images[[1]]$url,
                 stringsAsFactors = F) %>%
        mutate(album_release_date = GET(paste0('https://api.spotify.com/v1/albums/', str_replace(tmp$uri, 'spotify:album:', ''))) %>% content %>% .$release_date, # you need a separate call to on "albums" to get release date.
               album_release_year = ifelse(nchar(album_release_date) == 4, year(as.Date(album_release_date, '%Y')), year(as.Date(album_release_date, '%Y-%m-%d'))) # not all album_release_dates have months, so I created album_release year for sorting
        )
    } else {
      NULL
    }
    
  }) %>% filter(!duplicated(tolower(album_name))) %>%  # Sometimes there are multiple versions (just with different capitalizations) of the same album
    arrange(album_release_year) %>% 
   mutate(artist=artistData()$artist_name[j])



  if(i!=1) {
    allAlbum_info <- rbind(allAlbum_info,album_info)
  } else {
     allAlbum_info <- album_info
  }
 }  
 
    info=list(allAlbum_info=allAlbum_info)

  return(info)
 
 })
 
 
  



output$albumTable <- DT::renderDataTable({
  allArtistsData()$allAlbum_info %>%
    select(album=album_name) %>% 
                         DT::datatable(class='compact stripe hover row-border order-column',
                                       selection="multiple",rownames=FALSE,options= list(paging = FALSE, searching = FALSE,info=FALSE))
})

DT::dataTableOutput("albumTable")
actionButton(inputId="goAlbums",label="Get Albums")

```


You may note that often not all albums from a specific artist are available. I'm guessing that reflects what is available on the music service

If you want to EXCLUDE specific albums (e.g repeated compilations) then select them prior to pushing the button. The output will take a few seconds to arrive

---

```{r tracks, warning=FALSE}
# Get Tracks

  
   tracksData <- eventReactive(input$goAlbums,{

     
     selectedAlbums <- allArtistsData()$allAlbum_info$album_uri
     
     # remove any albums not requireed in further processing
     if (length(input$albumTable_rows_selected)>0) {
       
       t = as.integer(input$albumTable_rows_selected)
        remove <- selectedAlbums[t]
  selectedAlbums <- setdiff(selectedAlbums,remove)
     }

    track_info <- map_df(selectedAlbums, function(x) {
    tracks <- GET(paste0('https://api.spotify.com/v1/albums/', x, '/tracks')) %>%
      content %>%
      .$items
     
    uris <- map(1:length(tracks), function(z) {
      gsub('spotify:track:', '', tracks[z][[1]]$uri)
    }) %>% unlist %>% paste0(collapse=',')
     
    res <- GET(paste0('https://api.spotify.com/v1/audio-features/?ids=', uris),
               query = list(access_token = access_token)) %>% content %>% .$audio_features
    df <- unlist(res) %>%
      matrix(nrow = length(res), byrow = T) %>%
      as.data.frame(stringsAsFactors = F)
    names(df) <- names(res[[1]])
    df <- df %>%
      mutate(album_uri = x,
             track_number = row_number()) %>%
      rowwise %>%
      mutate(track_name = tracks[[track_number]]$name) %>%
      ungroup %>%
      #left_join(album_info, by = 'album_uri') %>%
      left_join(allArtistsData()$allAlbum_info, by = 'album_uri') %>%
      rename(track_uri = id) %>%
      select(-c(type, track_href, analysis_url, uri))
    return(df) ## works to here
  }) %>%
#  mutate(artist_img = artist_info$artist_img) %>%
#       mutate(artist_img =  artistData()$artist_img) %>% removed , 'artist_img' from list below - could add [1]?
    mutate_at(c('album_uri', 'track_uri', 'album_release_date', 'track_name', 'album_name'), funs(as.character)) %>%
    mutate_at(c('danceability', 'energy', 'key', 'loudness', 'mode', 'speechiness', 'acousticness', 'album_release_year',
                'instrumentalness', 'liveness', 'valence', 'tempo', 'duration_ms', 'time_signature', 'track_number'), funs(as.numeric(gsub('[^0-9.-]+', '', as.character(.))))) # for some reason parse_number() from readr doesn't work here
     write_csv(track_info,"beyonceTrackInfo.csv")
  return(track_info)
   })

 
 output$tracksTable <- DT::renderDataTable({
  tracksData() %>%
     select(artist,album=album_name,release=album_release_year, track=track_name,danceability,instrumentalness,valence)  %>% 
                         DT::datatable(class='compact stripe hover row-border order-column',
                                       rownames=FALSE,options= list(paging = TRUE, searching = TRUE,info=FALSE))
})

DT::dataTableOutput("tracksTable")

```

---

Firstly, you will see a searchable, sortable table of all the tracks. Included are three of Spotify's metrics
"Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry)"


---

Switch between the audio features, as desired. 


```{r, warning=FALSE}

radioButtons(inputId="metric",label=NULL,choices=c("Danceability","Instrumentalness","Valence"), inline=TRUE)
hr()

 output$tracksPlot <- renderPlotly({
   # set for interactivity
   nms <- row.names(tracksData())
 
 #  choice <- "Danceability"
  
   if (input$metric== "Danceability") {
  p <-  tracksData() %>%
    filter(danceability>=0) %>% 
       plot_ly(x=~album_release_date,y=~danceability, key= nms, source = "tracks") %>% 
       add_markers(hoverinfo="text", name="The Artist",
              text=~paste0(album_name,
                           "<br>",track_name,
                           "<br>",danceability))
   } else if (input$metric== "Instrumentalnes") {
  p <-  tracksData() %>%
    filter(instrumentalnes>=0) %>% 
       plot_ly(x=~album_release_date,y=~instrumentalnes, key= nms, source = "tracks") %>% 
       add_markers(hoverinfo="text", name="The Artist",
              text=~paste0(album_name,
                           "<br>",track_name,
                           "<br>",instrumentalness))
   } else {
  p <-  tracksData() %>%
    filter(valence>=0) %>% 
       plot_ly(x=~album_release_date,y=~valence, key= nms, source = "tracks") %>% 
       add_markers(hoverinfo="text", name="The Artist",
              text=~paste0(album_name,
                           "<br>",track_name,
                           "<br>",valence))
   } 
  p %>% 
  layout(title=paste0(input$metric,"<br>Hover for Info. Click for Video"),
         xaxis=list(title="",showticklabels=FALSE),
         yaxis=list(title=input$metric))%>%
  config(displayModeBar = F, showLink = F)
})

plotlyOutput("tracksPlot")
 

```



The albums are ordered by release date. Zoom in, if desired; hover points to view details and click to watch the first listed version on youtube, if available. Full screening is an option




```{r video, warning=FALSE}



output$video <- renderUI({
  
  # retrieve the clicked point, translate to artist/song combo
  d <- event_data("plotly_click", source = "tracks")
   s <- as.integer(d$key)

  song  <- paste(tracksData()$artist[s],tracksData()$track_name[s],sep="+")
  song  <- str_replace_all(song," ","+")
  
  
 # navigate to correct index page on youtube and select first video link 
   ytURL <- paste0("https://www.youtube.com/results?search_query=",song)
 
 
 ytHTML <- read_html(ytURL)
 link <- ytHTML %>% # dpwnloads all the html
   html_nodes( "h3 a" ) %>%
   html_attr("href")
 

 if (length(link)>0) {
   theLink <- link[[1]]
 } else {
   theLink <- NA
 }
  theLink <- str_split(theLink,"=")[[1]][2]

  # using the vembedr package embed the resultant video
  embed_youtube(id = theLink,width = 400)
  })

  uiOutput("video")


```

For many songs, the lyric is also available

```{r}

output$lyric <- renderUI({
  
  # obtain song title from plot click
  d <- event_data("plotly_click", source = "tracks")
   s <- as.integer(d$key)
 
  song  <- tracksData()$track_name[s]
 

  
songUrl <- str_replace_all(song," ","%20")

lyricURL <- paste0("https://www.google.ca/search?q=",songUrl,"%20lyrics&rct=j")

# scrapes appropriate url for all hrefs 
urls <- read_html(lyricURL) %>% 
  html_nodes("a") %>% 
  html_attr("href")


# Discover which is relevant url out of many
x <- which(map_lgl(urls, ~ str_sub(.x,1,29)=="https://play.google.com/music")==TRUE) 

# This attempts to cater for instances where title has an addition e.g "- Ramastered""
if (length(x)==0) {
  
  song  <- str_trim(str_split(song,"-")[[1]][1])

  
songUrl <- str_replace_all(song," ","%20")

lyricURL <- paste0("https://www.google.ca/search?q=",songUrl,"%20lyrics&rct=j")

urls <- read_html(lyricURL) %>% 
  html_nodes("a") %>% 
  html_attr("href")

x <- which(map_lgl(urls, ~ str_sub(.x,1,29)=="https://play.google.com/music")==TRUE)
  
}

# now obtain the full lyric from Google Music - or print failure
if (length(x)>0) {
fullLyric <- urls[x]

outcome <-fullLyric %>% 
  read_html() %>% 
  html_nodes('p') %>% 
  paste(collapse = '\n')
} else {
  outcome <- "No Lyrics on Google Music available"
}

 
HTML(outcome)
  
})
br()
uiOutput("lyric")


```